---
title: "10x PAH NucSeq - CellChat Immune vs Endothelial+Stromal"
author: "Yvon Mbouamboua<BR>mbouamboua@ipmc.cnrs.fr"
date: 'Compiled: `r Sys.Date()`'
output:
  rmdformats::readthedown:
    self-contained: yes
    highlight: haddock
    #css: styles.css
    theme: flatly
    number_sections: yes
    thumbnails: no
  html_notebook:
    theme: cerulean 
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---


# Setting parameters

```{r params}

# Chunck options
knitr::opts_chunk$set(echo = T,cache = F,cache.lazy = F,tidy = T,message = F,warning = F)
options(future.globals.maxSize = 80000*1024^2)
main.dir <- "/data/data_mbouamboua/projects/10x_htap_nucseq/"
out.dir <- paste0(main.dir, "github/007_cellchat_immune_vs_endo_stroma/")
if(!dir.exists(out.dir)){dir.create(file.path(out.dir), recursive = T, showWarnings = F)}
set.seed(4743)


```

# Packages and sources

```{r Libs, message=F}

library(Matrix)
library(dplyr)
library(Seurat)
library(scater)
library(ggplot2)
library(patchwork)
library(NMF)
library(CellChat)
library(ggalluvial)
library(ComplexHeatmap)

#source("/data/data_mbouamboua/apps/Rpkgs/Rpkgs.R")
source("/data/data_mbouamboua/tools/utilities.R")
source("/data/data_mbouamboua/tools/visualization.R")
#source("/data/data_mbouamboua/tools/differential_expression.R")

```


# Define colors

```{r}

Endothelial <- c("Lymph","aCap","gCap","Art","V-Pulm","V-Sys")
Immune <- c("AM1", "AM2","AM-prolif","DC1","DC2","Mono","Inter-Macro","CD4","CD8","T-prolif","NK","Mast","Plasma","pDC","B","Megak")
Stroma <- c("Peri", "ASMC","VSMC","AdvFibro","AlvFibro", "MyoFibro")

col.pal <- list()
col.pal$group <- setNames(c("#006E82","#AA0A3C"), c("CTRL", "PAH"))

col.pal$endo <-  setNames(c(cartography::carto.pal(pal1 = "blue.pal", n1 = 6)),Endothelial)
col.pal$stroma <-  setNames(c(cartography::carto.pal(pal1 = "wine.pal", n1 = 6)),Stroma)
col.pal$imm <-  setNames(c(cartography::carto.pal(pal1 = "green.pal", n1 = 7),
                                           cartography::carto.pal(pal1 = "kaki.pal", n1 = 4),
                                           cartography::carto.pal(pal1 = "brown.pal", n1 = 5)),Immune)
col.pal$celltype <- c(col.pal$endo, col.pal$imm,col.pal$stroma)

col.order <-  c("#83BED9","#CD4A50","#AE282C","#82B577","#BEDCB0","#A0C893","#3D87AF","#E69498","#633713",
                "#A8AB8E","#858A61","#5F9E55","#37842E","#65A6C7","#155F28", "#A2D6EC","#D1B695","#23120A",
                "#217024","#781411","#353C20","#8F6843","#F2B4B7","#B2916E","#5A632A","#1F6A95","#115379","#DB7479")

```



## Load cellchat data

```{r message=FALSE, warning=FALSE}

cellchat.CTRL <- readRDS(paste0(main.dir, "github/002_data_objects/cellchat.CTRL.imm.endo.stroma.rds"))
cellchat.PAH <- readRDS(paste0(main.dir, "github/002_data_objects/cellchat.PAH.imm.endo.stroma.rds"))
object.list <- list(CTRL = cellchat.CTRL, PAH = cellchat.PAH)
cellchat <- mergeCellChat(object.list, add.names = names(object.list))

```



```{r warnning = F, fig.width=3, fig.height=6}

rankNet(cellchat,
        mode = "comparison", 
        stacked = T, 
        do.stat = T, 
        do.flip = T,
        color.use = col.pal$group) +
  theme(axis.text.x = element_text(size = 8, color = "black", hjust = 0.5),
        axis.text.y = element_text(face = "bold", size = 8),
        axis.title = element_text(size = 8),
        legend.position = "none",
        legend.key.size = unit(0.3, "cm"))

ggsave(paste0(out.dir, "rankNet.pdf"), width = 3, height = 6, dpi = 100, useDingbats = F)

```

## Compare outgoing (or incoming) signaling associated with each cell population

```{r fig.width=12, fig.height=10}

i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)

ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 10, height = 14, color.heatmap = "OrRd", color.use = col.order)

ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i+1], width = 10, height = 14, color.heatmap = "OrRd",color.use = col.order)

draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
pdf(paste0(out.dir, "outgoing.singaling.pattern.pdf"), width = 11, height = 11, useDingbats = F)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
dev.off()

i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht3 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i], width = 10, height = 14, color.heatmap = "OrRd",color.use = col.order)
ht4 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i+1], width = 10, height = 14, color.heatmap = "OrRd", color.use = col.order)
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))
pdf(paste0(out.dir,"incoming.singaling.pattern.pdf"), width = 11, height = 11, useDingbats = F)
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))
dev.off()

i = 1
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathway.union, title = names(object.list)[i], width = 10, height = 14, color.heatmap = "OrRd", color.use = colors)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathway.union, title = names(object.list)[i+1], width = 10, height = 14, color.heatmap = "OrRd", color.use = col.order)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
pdf(paste0(out.dir, "overall.singaling.pattern.pdf"), width = 11, height = 11, useDingbats = F)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
dev.off()

```


# Differential number of interactions or interaction strength among different cell populations

```{r fig.width=10, fig.height=6}

sources.use <- c("AM1", "AM2","AM-prolif","DC1","DC2","Mono","Inter-Macro","CD4","CD8","T-prolif","NK","Mast","Plasma","pDC","B","Megak")
targets.use <- c("Lymph","aCap","gCap","Art","V-Pulm","V-Sys","Peri", "ASMC","VSMC","AdvFibro","AlvFibro", "MyoFibro")
# Do heatmap based on a merged object
gg1 <- netVisual_heatmap(cellchat, color.use = col.order)
gg2 <- netVisual_heatmap(cellchat, color.use = col.order, measure = "weight")
p <- gg1 + gg2

#pdf(paste0(out.dir, "/04_data_objects/cellchat/netVisual_interaction_heatmap.pdf"), width = 10, height = 6, useDingbats = F)
print(p)
#dev.off()

```






# LR pairs

```{r fig.width=10, fig.height=5}

invisible(gc())
dir <- paste0(out.dir, "LR.pairs/")
if(!dir.exists(dir)){dir.create(file.path(dir), recursive = T, showWarnings = F)}

joined <- data.frame()
celltype.to.test <- as.vector(unique(cellchat@idents[[1]]))
#celltype.to.test <- celltype.to.test[!celltype.to.test %in% c("PNEC","AM-prolif","T-prolif")]


for(i in 1:length(celltype.to.test)){
  print(celltype.to.test[i])
  netVisual <- netVisual_bubble(cellchat,
                                sources.use = celltype.to.test[i],
                                remove.isolate = T,
                                targets.use = c(celltype.to.test),
                                comparison = c(1, 2),
                                angle.x = 45,
                                color.heatmap = "Spectral",
                                color.text = col.pal$group,
                                title.name = paste("LR pairs -", celltype.to.test[i]),
                                return.data = F
  ) +
    theme(text = element_text(size = 8),
          legend.key.height = unit(0.2, 'cm'),
          legend.key.width = unit(0.2, 'cm')) +
    guides(colour = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
    geom_point(shape = 21, colour="black", stroke=0.5) 
  
  pdf(paste(dir, celltype.to.test[i], "_LR_pairs.pdf", sep = ""), width = 10, height = 5, useDingbats = F)
  print(netVisual)
  dev.off()
  netVisual <- netVisual$data
  rownames(netVisual) <- c()
  joined <- rbind(joined, netVisual)
}

write.table(joined, file = paste(dir, "LR.pairs.tsv", sep = "/"), sep = "\t", quote = F, row.names = F, col.names = T)

```



# Contribution of each L-R pairs

```{r fig.width=4.5, fig.height=3}

dir <- paste0(out.dir, "LR.pair_contributions/")
if(!dir.exists(dir)){dir.create(file.path(dir), recursive = T, showWarnings = F)}
LR.pairs <- read.delim(paste0(out.dir, "LR.pairs/LR.pairs.tsv"))
pathway.name <- unique(LR.pairs$pathway_name)
ranknet <- rankNet(cellchat, mode = "comparison")
ranknet <- ranknet$data
only.CTRL.pathway.name <- subset(ranknet, contribution == 0 & group == "PAH")
only.CTRL.pathway.name <- as.vector(only.CTRL.pathway.name$name)
only.PAH.pathway.name <- subset(ranknet, contribution == 0 & group == "CTRL" ) 
only.PAH.pathway.name <- as.vector(only.PAH.pathway.name$name)
pathway.name <- unique(LR.pairs$pathway_name)
pathway.name <-  pathway.name[!pathway.name %in% c(only.CTRL.pathway.name, only.PAH.pathway.name)]


for (pathway in 1:length(pathway.name)) {
  pathways.show <- pathway.name[pathway]
  for (i in 1:length(object.list)) {
    netAnalysis_contribution(object.list[[i]],
                             signaling = pathways.show,
                             title = "") +
      theme(
            plot.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text.y = element_text(size = 10))
    
    ggsave(paste(dir, paste0(pathways.show,"_",names(object.list)[i]),".pdf"), 
        width = 3.5, height = 3, useDingbats = F)
  }
}


# only.PAH.pathway.name
for (pathway in 1:length(only.PAH.pathway.name)) {
  pathways.show <- only.PAH.pathway.name[pathway]
    netAnalysis_contribution(object = cellchat.PAH,
                             signaling = pathways.show,
                             title = "") +
      theme(plot.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text.y = element_text(size = 10))
    
    ggsave(paste0(dir, pathways.show,"_PAH.pdf"), 
        width = 3.5, height = 3, useDingbats = F)
}


# only.CTRL.pathway.name
for (pathway in 1:length(only.CTRL.pathway.name)) {
  pathways.show <- only.CTRL.pathway.name[pathway]
    netAnalysis_contribution(object = cellchat.CTRL,
                             signaling = pathways.show,
                             title = "") +
      theme(plot.title = element_text(size = 10),
            axis.title = element_text(size = 10),
            axis.text.y = element_text(size = 10))
    
    ggsave(paste0(dir, pathways.show,"_CTRL.pdf"), 
        width = 3.5, height = 3, useDingbats = F)
}


```




# Subset BMP+TGFb LR pairs

```{r }

dir <- paste0(out.dir, "BMP_TGFb_LR.pairs/")
if(!dir.exists(dir)){dir.create(file.path(dir), recursive = T, showWarnings = F)}

# BMP
LR.pairs <- read.delim(paste0(out.dir, "LR.pairs/LR.pairs.tsv"))
#pathway.name <- unique(LR.pairs$pathway_name)
pathways.show <- c("BMP")

pairLR.use <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = F, enriched.only = T)
LR.pairs.sub <- subset(LR.pairs, pathway_name == paste(pathways.show))
sources.use <- unique(LR.pairs.sub$source)
targets.use <- unique(LR.pairs.sub$target)

for(i in 1:length(sources.use)){

  plot <- netVisual_bubble(cellchat,
                           pairLR.use = pairLR.use,
                           sources.use = sources.use[i],
                           targets.use = targets.use,
                           remove.isolate = T,
                           comparison = c(1, 2),
                           angle.x = 45,
                           color.heatmap = "Spectral",
                           color.text = col.pal$group,
                           title.name = paste(pathways.show, " LR pairs - ", sources.use[i]),
                           return.data = F
  ) +
    theme(text = element_text(size = 8),
          legend.key.height = unit(0.2, 'cm'),
          legend.key.width = unit(0.2, 'cm')) +
    guides(colour = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
    geom_point(shape = 21, colour="black", stroke=0.5)
  
  pdf(paste(dir, sources.use[i], "_BMP_LR_pairs.pdf", sep = ""), width = 7, height = 3, useDingbats = F)
  print(plot)
  dev.off()
}



# TGFb
LR.pairs <- read.delim(paste0(out.dir, "LR.pairs/LR.pairs.tsv"))
#pathway.name <- unique(LR.pairs$pathway_name)
pathways.show <- c("TGFb")

pairLR.use <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = F, enriched.only = T)
LR.pairs.sub <- subset(LR.pairs, pathway_name == paste(pathways.show))
sources.use <- unique(LR.pairs.sub$source)
targets.use <- unique(LR.pairs.sub$target)

for(i in 1:length(sources.use)){

  plot <- netVisual_bubble(cellchat,
                           pairLR.use = pairLR.use,
                           sources.use = sources.use[i],
                           targets.use = targets.use,
                           remove.isolate = T,
                           comparison = c(1, 2),
                           angle.x = 45,
                           color.heatmap = "Spectral",
                           color.text = col.pal$group,
                           title.name = paste(pathways.show, " LR pairs - ", sources.use[i]),
                           return.data = F
  ) +
    theme(text = element_text(size = 8),
          legend.key.height = unit(0.2, 'cm'),
          legend.key.width = unit(0.2, 'cm')) +
    guides(colour = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
    geom_point(shape = 21, colour="black", stroke=0.5)
  
  pdf(paste(dir, sources.use[i], "_TGFb_LR_pairs.pdf", sep = ""), width = 7, height = 3, useDingbats = F)
  print(plot)
  dev.off()
}



```




# Signaling network

```{r}

dir <- paste0(out.dir, "signaling_network/")
if(!dir.exists(dir)){dir.create(file.path(dir), recursive = T, showWarnings = F)}

LR.pairs <- read.delim(paste0(out.dir, "LR.pairs/LR.pairs.tsv"))
ranknet <- rankNet(cellchat, mode = "comparison")
ranknet <- ranknet$data
only.CTRL.pathway.name <- subset(ranknet, contribution == 0 & group == "PAH")
only.CTRL.pathway.name <- as.vector(only.CTRL.pathway.name$name)
only.PAH.pathway.name <- subset(ranknet, contribution == 0 & group == "CTRL" ) 
only.PAH.pathway.name <- as.vector(only.PAH.pathway.name$name)
pathway.name <- unique(LR.pairs$pathway_name)
pathway.name <-  pathway.name[!pathway.name %in% c(only.CTRL.pathway.name, only.PAH.pathway.name)]

for (pathway in 1:length(pathway.name)) {
  
  pathways.show <- pathway.name[pathway]
  weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathways.show)
  
  for (i in 1:length(object.list)) {
    pdf(paste(dir, paste0(pathways.show,"_",names(object.list)[i]),".pdf", sep = ""), width = 7, height = 7, useDingbats = F)
    netVisual_aggregate(object.list[[i]], 
                        signaling = pathways.show, 
                        remove.isolate = F,
                        layout = "chord", 
                        #sources.use = sources.use, 
                        #targets.use = targets.use,
                        color.use = col.pal$celltype,
                        edge.weight.max = weight.max[1], 
                        edge.width.max = 10, 
                        signaling.name = paste(pathways.show, names(object.list)[i]))
    dev.off()
  }
}


# only.PAH.pathway.name

for (pathway in 1:length(only.PAH.pathway.name)) {
  
  pathways.show <- only.PAH.pathway.name[pathway]

  pdf(paste0(dir, pathways.show,"_PAH.pdf"), width = 7, height = 7, useDingbats = F)

  netVisual_aggregate(cellchat.PAH, 
                      signaling = pathways.show, 
                      remove.isolate = F,
                      layout = "chord", 
                      color.use = col.pal$celltype,
                      edge.width.max = 10, 
                      signaling.name = paste(pathways.show, " PAH")
  )
  dev.off()
}


# only.CTRL.pathway.name

for (pathway in 1:length(only.CTRL.pathway.name)) {
  
  pathways.show <- only.CTRL.pathway.name[pathway]

  pdf(paste0(dir, pathways.show,"_CTRL.pdf"), width = 7, height = 7, useDingbats = F)
  
  netVisual_aggregate(cellchat.CTRL, 
                      signaling = pathways.show, 
                      remove.isolate = F,
                      layout = "chord", 
                      color.use = col.pal$celltype,
                      edge.width.max = 10, 
                      signaling.name = paste(pathways.show, " CTRL")
  )
  dev.off()
}

```



# Signalig role

```{r}

# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
cellchat.CTRL <- netAnalysis_computeCentrality(cellchat.CTRL)
gg1 <- netAnalysis_signalingRole_scatter(cellchat.CTRL)
# Signaling role analysis on the cell-cell communication networks of interest
gg2 <- netAnalysis_signalingRole_scatter(cellchat.CTRL, signaling = c("BMP"), color.use = col.pal$celltype)
gg1 + gg2

```


# Pathway signaling roles

```{r fig.width=8, fig.height=4}

#cellchat.CTRL <- netAnalysis_computeCentrality(cellchat.CTRL, slot.name = "netP")
#pathways.show <- c("ACTIVIN","BMP", "CCL","CXCL","EDN","GDF", "ncWNT","NRG", "PERIOSTIN","SEMA3", "TRAIL", "VGF", "VISFATIN", "WNT")

invisible(gc())
dir <- paste(out.dir, "pathway_signaling_roles/CTRL/", sep = "/")
if(!dir.exists(dir)){dir.create(file.path(dir), recursive = TRUE, showWarnings = FALSE)}

pathways.show <- cellchat.CTRL@netP$pathways

for(i in 1:length(pathways.show)){
  print(pathways.show[i])
  pdf(paste(dir,  pathways.show[i], ".pdf", sep = ""), width = 8, height = 2.5, useDingbats = FALSE)
  netAnalysis_signalingRole_network(cellchat.CTRL, signaling = pathways.show[i],  width = 12,
                                    height = 2.5, font.size = 7, color.use = col.order) 
  dev.off()
  }

#cellchat.PAH <- netAnalysis_computeCentrality(cellchat.PAH, slot.name = "netP")

invisible(gc())
dir <- paste(out.dir, "pathway_signaling_roles/PAH/", sep = "/")
if(!dir.exists(dir)){dir.create(file.path(dir), recursive = TRUE, showWarnings = FALSE)}

pathways.show <- cellchat.PAH@netP$pathways

for(i in 1:length(pathways.show)){
  print(pathways.show[i])
  pdf(paste(dir,  pathways.show[i], ".pdf", sep = ""), width = 8, height = 2.5, useDingbats = FALSE)
  netAnalysis_signalingRole_network(cellchat.PAH, signaling = pathways.show[i],  width = 12, 
                                    height = 2.5, font.size = 7, color.use = col.order) 
  dev.off()
}

```



## Celltype signaling

```{r fig.width=10, fig.height=10}

## CTRL celltype signaling
groupize <- as.numeric(table(cellchat.CTRL@idents)) # number of cells in each cell group
mat <- cellchat.CTRL@net$weight
#par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  p <- netVisual_circle(mat2, vertex.weight = groupize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i], color.use = col.order, vertex.label.cex = 0.7)
  pdf(paste(out.dir, "celltype_signaling_roles/CTRL/", rownames(mat)[i], "_CTRL.pdf", sep = ""), width = 8, height = 8, useDingbats = FALSE)
  print(p)
  dev.off()
}

## PAH celltype signaling
groupize <- as.numeric(table(cellchat.PAH@idents)) # number of cells in each cell group
mat <- cellchat.PAH@net$weight
#par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  p <- netVisual_circle(mat2, vertex.weight = groupize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i], color.use = col.order, vertex.label.cex = 0.7)
  pdf(paste(out.dir, "celltype_signaling_roles/PAH/", rownames(mat)[i], "_PAH.pdf", sep = ""), width = 8, height = 8, useDingbats = FALSE)
  print(p)
  dev.off()
}

```




# plotGeneExpression

```{r warning=F, message=F, fig.width=4, fig.height=4}

sobj <- readRDS(paste0(main.dir, "github/002_data_objects/htap.filt.rds"))
Idents(sobj) <- "population"
sobj <- subset(sobj, idents = c("Endothelial", "Stroma", "Immune"))

dir <- paste0(out.dir, "pathway_gene_expression/")
if(!dir.exists(dir)){dir.create(file.path(dir), recursive = T, showWarnings = F)}

LR.pairs <- read.delim(paste0(out.dir, "LR.pairs/LR.pairs.tsv"))
#pairLR.use <- extractEnrichedLR(cellchat.CTRL, signaling = pathways.show, geneLR.return = F)
pathway.name <- unique(LR.pairs$pathway_name)
DefaultAssay(sobj) <- "RNA"

for (pathway in 1:length(pathway.name)) {
  
  pathways.show <- pathway.name[pathway]
  LR.pairs.sub <- subset(LR.pairs, pathway_name == paste(pathways.show))
  ligand <- unique(LR.pairs.sub$ligand)
  ligand <- unique(unlist(strsplit(x = ligand, "\\_")))
  receptor <- unique(LR.pairs.sub$receptor)
  receptor <- unique(unlist(strsplit(x = receptor, "\\_")))
  genes <- unique(c(ligand, receptor))
  genes <- rownames(sobj[genes,])
  genes <- toupper(genes)
  
  StackedVlnPlot(obj = sobj, angle.x = 45,
                 features = genes,
                 group.by = "celltype_order", 
                 split.by = "group", 
                 color.text.x = col.pal$celltype,
                 color.use = col.pal$group)
  
   ggsave(paste0(dir, pathways.show,"_gene_expression.pdf"), 
        width = 10, height = 4, useDingbats = F)

}


```

