---
title: "10x PAH NucSeq - Gen Set Enrichment Analysis"
author: '[Yvon Mbouamboua](mbouamboua@ipmc.cnrs.fr)'
output:
  html_document: default
  html_notebook: default
date: 'Compiled: `r Sys.Date()`'
---

# Setting parameters

```{r setup, include=T, warning=FALSE}

options(future.globals.maxSize = 80000*1024^2)
main.dir <- "/data/data_mbouamboua/projects/10x_htap_nucseq/"
out.dir <- paste0(main.dir, "004_outs/GSEA/")
if(!dir.exists(out.dir)){dir.create(file.path(out.dir), recursive = T, showWarnings = F)}
set.seed(4743)

```

# Require R packages and functions


```{r warning=F, message=F}

source("/data/data_mbouamboua/apps/Rpkgs/Rpkgs.R")
source("/data/data_mbouamboua/tools/utilities.R")
source("/data/data_mbouamboua/tools/visualization.R")
source("/data/data_mbouamboua/tools/differential_expression.R")
set.seed(4743)

```



# Define function

```{r fig.width=8, fig.height=8}

# Adapted from: https://github.com/xmc811/Scillus/blob/master/R/visualization.R

Get_GO <- function(markers, 
                   #cluster_name = NULL, 
                   topn = 100, 
                   org = "human", 
                   ont = "BP", 
                   category,
                   out.dir,
                   out.name,...) {
  
  pkg_name <- ifelse(org == "human", "org.Hs.eg.db", "org.Mm.eg.db")
  
  if (!requireNamespace(pkg_name, quietly = TRUE)) {
    stop(paste("Package", pkg_name, "needed for this function to work. Please install it."),
         call. = FALSE)
  }
  
  if (!requireNamespace("clusterProfiler", quietly = TRUE)) {
    stop(paste("Package \"clusterProfiler\" needed for this function to work. Please install it."),
         call. = FALSE)
  }
  
  
  joined <- data.frame()
  cluster_names <- as.vector(unique(markers$cluster))
  
  for(i in 1:length(cluster_names)){
    cluster_name <- cluster_names[i]
    
    gene_list <- markers %>% 
      filter(.data$cluster == cluster_name) %>% 
      arrange(.data$p_val_adj) %>% 
      head(topn) %>% 
      pull(.data$gene)
    
    db <- if(org == "human") org.Hs.eg.db::org.Hs.eg.db else org.Mm.eg.db::org.Mm.eg.db
    
    res <- clusterProfiler::enrichGO(gene = gene_list, 
                                     OrgDb = db, 
                                     ont = ont,
                                     keyType = "SYMBOL", ...)
    
    dir <- paste(out.dir, ont, out.name, "/", sep = "/")
    if(!dir.exists(dir)){
      dir.create(file.path(dir), recursive = TRUE, showWarnings = FALSE)
    }
    
    res@result$cluster_name <- replicate(nrow(res@result),paste(cluster_name))
    saveRDS(res, paste(dir, cluster_name, "_GO_", ont, ".rds", sep = ""))
    
    df <- res@result
    df$cluster_name <- replicate(nrow(df),paste(cluster_name))
    rownames(df) <- c()
    joined <- rbind(joined, df)

    df <- as_tibble(res@result) %>% 
      arrange(.data$p.adjust) %>% 
      head(category) %>% 
      mutate(cluster = cluster_name) %>% 
      mutate(Description = stringr::str_to_title(.data$Description)) %>% 
      mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))
    
      #df <- df[grep("^Epithel", df$Description, invert = TRUE),]
      #df <- df %>% dplyr::filter(!grepl('^epithel', Description))
      
    ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
      geom_bar(aes(fill = .data$Count), 
               stat = "identity") + 
      scale_fill_gradientn("Gene Count", 
                           colours = colorspace::diverge_hcl(7), 
                           guide = "colourbar") +
      coord_flip() + 
      #geom_hline(yintercept = -log10(0.05), linetype = "dashed") + 
      xlab("Gene Ontology") + 
      ylab(bquote("-log"[10] ~ " adjusted p-value")) +
      theme_bw() +
      theme(axis.text = element_text(size = 8, color = "black"),
            axis.title = element_text(size = 10, color = "black"),
            legend.title=element_text(size = 8, color = "black")) +
      ggtitle(cluster_name)
    
    saveRDS(res, paste(dir, cluster_name, "_GO_", ont, ".rds", sep = ""))
    
    ggsave(paste(dir, cluster_name, "_GO_", ont, ".pdf", sep = ""), width = 8, height = 8)
    
  }
  
  write.table(joined, file = paste(dir, "all_res_GO_", ont, ".tsv", sep = ""),
              sep = "\t", quote = F, row.names = F, col.names = T)
  
}

```


# All celltypes

```{r fig.width=2.5, fig.height=4}

markers <- read.delim(paste0(main.dir, "004_outs/005_DE_markers/celltype_markers.tsv"), sep = "\t")
markers <- markers %>% dplyr::filter(p_val_adj <= 0.05, avg_log2FC > 0)

Get_GO(markers = markers, 
                topn = 10000, 
                category = 50, 
                org = "human", 
                ont = "BP", 
                out.dir = out.dir,
                out.name = "GSEA_NucSeq")

```



# Epithelial cells
## Get gene markers



## Get GO BP
```{r fig.width=2.5, fig.height=4}

markers <- read.delim(paste0(main.dir, "004_outs/005_DE_markers/epi_markers.tsv"), sep = "\t")
markers <- markers %>% dplyr::filter(p_val_adj <= 0.05, avg_log2FC > 0)

## GO BP
Get_GO(markers = markers, 
                topn = 10000, 
                category = 50, 
                org = "human", 
                ont = "BP", 
                out.dir = out.dir,
                out.name = "Epithelial")

```


 


# Endothelial cells

## Get gene markers

```{r fig.width=2.5, fig.height=11}

invisible(gc())
sobj <- readRDS(paste0(main.dir, "github/002_data_objects/endothelial.rds"))
markers <- Get_Markers(obj = sobj, group.by = "celltype", test_use = "wilcox",only.pos = F)
head(markers)
write.table(markers, paste(out.dir, "endo_markers.tsv", sep = "/"), sep = "\t", row.names = FALSE, quote = FALSE, col.names = TRUE)
# top <- markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_log2FC)
# Idents(sobj) <- "celltype"
# Dot_Plot(sobj = sobj, features = rev(top$gene), x.lab.rotate = T, flip.axes = T, base.size = 10, theme.bw = T)
# ggsave(paste(out.dir, "endo_markers.png", sep = "/"), width = 2.5, height = 11)

```


## Get GO BP without negative genes

```{r fig.width=2.5, fig.height=4}

markers <- read.delim(paste(out.dir, "endo_markers.tsv", sep = "/"), sep = "\t")
markers <- markers %>% dplyr::filter(p_val_adj <= 0.05, avg_log2FC > 0)

Get_GO(markers = markers, 
                topn = 10000, 
                category = 50, 
                org = "human", 
                ont = "BP", 
                out.dir = out.dir,
                out.name = "Endothelial")

```


## Get GO BP with negative genes

```{r fig.width=2.5, fig.height=4}

markers <- read.delim(paste0(out.dir, "endo_markers.tsv"), sep = "\t")
markers <- markers %>% dplyr::filter(p_val_adj <= 0.05)

Get_GO(markers = markers, topn = 10000,  category = 50,  org = "human", 
       ont = "BP",  out.dir = out.dir, out.name = "Endothelial_v2")

```


## Score and preranked 

```{r}


# GO_BP <- readRDS(paste(outdir,  "GO_BP.rds", sep = "/"))
# GO_BP@result$Description <- factor(gsub("/", " or ", GO_BP_ALK1_High_BMP9_vs_ALK1_High@result$Description))
# terms <- c("negative regulation of pathway-restricted SMAD protein phosphorylation", "blood vessel remodeling", "negative regulation of angiogenesis", "blood circulation", "regulation of cell growth", "regulation of blood pressure", "positive regulation of cell differentiation", "negative regulation of vasculature development" , "endothelial cell differentiation", "regulation of cell death")
# GO_BP <- dplyr::filter(GO_BP, Description %in% terms)


obj.list <- list()
files = list.files(path = paste0(out.dir, "BP/Endothelial_v2/"), pattern = '.rds', full.names = T, recursive = F)
obj.list = lapply(files, function (x) readRDS(x))


for (i in 1:length(obj.list)) {
  
  cluster_name <- unique(obj.list[[i]]@result$cluster_name)
  dir <- paste0(out.dir, "gsea_preranked/", cluster_name, "/", sep = "")
  if(!dir.exists(dir)){dir.create(file.path(dir), recursive = T, showWarnings = F)}

  for(i in 1:length(obj.list[[i]]@result$ID)){
    enrichplot::gseaplot2(obj.list[[i]],  geneSetID = i,  base_size = 8,
                          title = obj.list[[i]]@result$Description[[i]]) +
      theme(axis.text.y = element_text(color = "black", size = 8),
            axis.text.x = element_text(color = "black", size = 8),
            axis.title = element_text(color = "black", size = 8))
    
    ggsave(paste0(dir, obj.list[[i]]@result$Description[[i]], ".pdf", sep = ""))
    
  }
}


enrichplot::gseaplot2(obj.list[[1]]@result,  geneSetID = 1,  base_size = 8,
                          title = obj.list[[1]]@result$Description[[1]]) +
      theme(axis.text.y = element_text(color = "black", size = 8),
            axis.text.x = element_text(color = "black", size = 8),
            axis.title = element_text(color = "black", size = 8))


```



# Stromal cells

```{r fig.width=2.5, fig.height=4}

markers <- read.delim(paste0(main.dir, "004_outs/005_DE_markers/stromal_markers.tsv"), sep = "\t")
markers <- markers %>% dplyr::filter(p_val_adj <= 0.05, avg_log2FC > 0)

Get_GO(markers = markers, 
                topn = 10000, 
                category = 50, 
                org = "human", 
                ont = "BP", 
                out.dir = out.dir,
                out.name = "Stromal")

```


# Immune cells

```{r fig.width=2.5, fig.height=4}

# markers <- read.delim(paste0(main.dir, "004_outs/005_DE_markers/imm_markers.tsv"), sep = "\t")
# markers <- markers %>% dplyr::filter(p_val_adj <= 0.05, avg_log2FC > 0)
# 
# Get_GO(markers = markers, 
#                 topn = 10000, 
#                 category = 50, 
#                 org = "human", 
#                 ont = "BP", 
#                 out.dir = out.dir,
#                 out.name = "Immune")

```

# Merge GO BP results

```{r}

Epi_GO_BP <- read.delim(paste0(main.dir, "github/GSEA/BP/Epithelial/all_res_GO_BP.tsv"), sep = "\t")
Epi_GO_BP$population <- rep("Epithelial", nrow(Epi_GO_BP))

Endo_GO_BP <- read.delim(paste0(main.dir, "github/GSEA/BP/Endothelial/all_res_GO_BP.tsv"), sep = "\t")
Endo_GO_BP$population <- rep("Endothelial", nrow(Endo_GO_BP))

Stroma_GO_BP <- read.delim(paste0(main.dir, "github/GSEA/BP/Stromal/all_res_GO_BP.tsv"), sep = "\t")
Stroma_GO_BP$population <- rep("Stromal", nrow(Stroma_GO_BP))

Imm_GO_BP <- read.delim(paste0(main.dir, "github/GSEA/BP/Immune/all_res_GO_BP.tsv"), sep = "\t")
Imm_GO_BP$population <- rep("Endothelial", nrow(Imm_GO_BP))

GO_PB <- rbind(Epi_GO_BP,Endo_GO_BP,Stroma_GO_BP,Imm_GO_BP)

write.table(GO_PB,file=paste0(main.dir, "github/GSEA/BP/GO_PB.tsv"),sep="\t",quote=F,row.names=T)

```



# Biological theme comparison

```{r}

data(gcSample)
str(gcSample) 
ck <- compareCluster(geneCluster = gcSample, fun = enrichKEGG)
ck <- setReadable(ck, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
head(ck) 

```

```{r}

data(geneList, package="DOSE")
gene <- names(geneList)[abs(geneList) > 2]
# Entrez gene ID
head(gene)


mydf <- data.frame(Entrez=names(geneList), FC=geneList)
mydf <- mydf[abs(mydf$FC) > 1,]
mydf$group <- "upregulated"
mydf$group[mydf$FC < 0] <- "downregulated"

mydf$othergroup <- "A"
mydf$othergroup[abs(mydf$FC) > 2] <- "B"

formula_res <- compareCluster(Entrez~group, data=mydf, fun="enrichKEGG")
  
```



# Plot some GO terms
## aCap

```{r fig.width=7.3, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Endothelial/aCap_GO_BP.rds")
ids <- c("GO:0045766","GO:0071560","GO:2000351","GO:0034614","GO:0000302")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
          coord_flip() + labs(title = "aCap", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "aCap_top_5_go_bp.pdf"), useDingbats = F, width = 7.3, height = 1.8,  dpi = 320)

```


## gCap

```{r fig.width=6.3, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Endothelial/gCap_GO_BP.rds")
ids <- c("GO:0003018","GO:0045446","GO:0071559","GO:2001236","GO:0030510")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
          coord_flip() + labs(title = "gCap", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "gCap_top_5_go_bp.pdf"), useDingbats = F, width = 6.3, height = 1.8,  dpi = 320)

```


## VSMC

```{r fig.width=6.5, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Stromal/VSMC_GO_BP.rds")
ids <- c("GO:0006937","GO:1990874","GO:0014912","GO:0071559","GO:0001974")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
            coord_flip() + labs(title = "VSMC", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "VSMC_top_5_go_bp.pdf"), useDingbats = F, width = 6.5, height = 1.8,  dpi = 320)

```


## Peri

```{r fig.width=6, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Stromal/Peri_GO_BP.rds")
ids <- c("GO:0016358","GO:0043542","GO:0045446","GO:0071559","GO:0002690")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
          coord_flip() + labs(title = "Peri", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "Peri_top_5_go_bp.pdf"), useDingbats = F, width = 6, height = 1.8,  dpi = 320)

```


## AdvFibro

```{r fig.width=7, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Endothelial/aCap_GO_BP.rds")
ids <- c("GO:0030198","GO:0071560","GO:0048146","GO:0071773","GO:2001234")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
        coord_flip() + labs(title = "AdvFibro", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "AdvFibro_top_5_go_bp.pdf"), useDingbats = F, width = 7, height = 1.8,  dpi = 320)

```


## AlvFibro

```{r fig.width=6.5, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Endothelial/gCap_GO_BP.rds")
ids <- c("GO:0060485","GO:0071453","GO:0036293","GO:0070099","GO:0014910")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
      coord_flip() + labs(title = "AlvFibro", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "AlvFibro_top_5_go_bp.pdf"), useDingbats = F, width = 7, height = 1.8,  dpi = 320)


```

## MacroATP10A

```{r fig.width=6, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Immune/AM1_GO_BP.rds")
ids <- c("GO:0030258","GO:0006869","GO:0048771","GO:0006979","GO:0071559")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
    coord_flip() + labs(title = "MacroATP10A", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "MacroATP10A_top_5_go_bp.pdf"), useDingbats = F, width = 6, height = 1.8,  dpi = 320)

```



## MacroCD68

```{r fig.width=6, fig.height=1.8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Immune/AM2_GO_BP.rds")
ids <- c("GO:0016358","GO:0001819","GO:0045454","GO:0001974","GO:0002367")
df <- subset(res@result, ID %in% ids)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(stat = "identity", fill = "#208799") + 
  coord_flip() + labs(title = "MacroCD68", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "MacroCD68_top_5_go_bp.pdf"), useDingbats = F, width = 6, height = 1.8,  dpi = 320)

```
# Top 30 GO BP

## aCap

```{r fig.width=8, fig.height=8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Endothelial/aCap_GO_BP.rds")
sub <- subset(TOP_30_GO_BP,  cluster_name == "aCap")
df <- subset(res@result, ID %in% sub$ID)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(aes(fill = .data$Count), stat = "identity") + 
  scale_fill_gradient("Gene Count", low = "grey", high = "brown") +
  coord_flip() + labs(title = "aCap", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.position = c(0.5, 0.09),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "aCap_top_30_go_bp.pdf"), useDingbats = F, width = 8, height = 8,  dpi = 320)


```


## gCap

```{r fig.width=8, fig.height=8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Endothelial/gCap_GO_BP.rds")
sub <- subset(TOP_30_GO_BP,  cluster_name == "gCap")
df <- subset(res@result, ID %in% sub$ID)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(aes(fill = .data$Count), stat = "identity") + 
  scale_fill_gradient("Gene Count", low = "grey", high = "brown") +
  coord_flip() + labs(title = "gCap", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.position = c(0.5, 0.09),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "gCap_top_30_go_bp.pdf"), useDingbats = F, width = 8, height = 8,  dpi = 320)


```



## gCap

```{r fig.width=8, fig.height=8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Stromal/AlvFibro_GO_BP.rds")
sub <- subset(TOP_30_GO_BP,  cluster_name == "AlvFibro")
df <- subset(res@result, ID %in% sub$ID)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(aes(fill = .data$Count), stat = "identity") + 
  scale_fill_gradient("Gene Count", low = "grey", high = "brown") +
  coord_flip() + labs(title = "AlvFibro", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.position = c(0.5, 0.09),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "AlvFibro_top_30_go_bp.pdf"), useDingbats = F, width = 8, height = 8,  dpi = 320)


```



## AdvFibro

```{r fig.width=9, fig.height=8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Stromal/AdvFibro_GO_BP.rds")
sub <- subset(TOP_30_GO_BP,  cluster_name == "AdvFibro")
df <- subset(res@result, ID %in% sub$ID)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(aes(fill = .data$Count), stat = "identity") + 
  scale_fill_gradient("Gene Count", low = "grey", high = "brown") +
  coord_flip() + labs(title = "AdvFibro", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.position = c(0.5, 0.09),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "AdvFibro_top_30_go_bp.pdf"), useDingbats = F, width = 9, height = 8,  dpi = 320)


```




## Peri

```{r fig.width=8, fig.height=8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Stromal/Peri_GO_BP.rds")
sub <- subset(TOP_30_GO_BP,  cluster_name == "Peri")
df <- subset(res@result, ID %in% sub$ID)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(aes(fill = .data$Count), stat = "identity") + 
  scale_fill_gradient("Gene Count", low = "grey", high = "brown") +
  coord_flip() + labs(title = "Peri", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.position = c(0.5, 0.09),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "Peri_top_30_go_bp.pdf"), useDingbats = F, width = 8, height = 8,  dpi = 320)


```


## MacroCD68

```{r fig.width=8, fig.height=8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Immune/AM2_GO_BP.rds")
sub <- subset(TOP_30_GO_BP,  cluster_name == "MacroCD68")
df <- subset(res@result, ID %in% sub$ID)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(aes(fill = .data$Count), stat = "identity") + 
  scale_fill_gradient("Gene Count", low = "grey", high = "brown") +
  coord_flip() + labs(title = "MacroCD68", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.position = c(0.3, 0.09),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "MacroCD68_top_30_go_bp.pdf"), useDingbats = F, width = 8, height = 8,  dpi = 320)


```


## MacroATP10A


```{r fig.width=8, fig.height=8}

res <- readRDS("/data/data_mbouamboua/projects/10x_htap_nucseq/004_outs/GSEA/BP/Immune/AM1_GO_BP.rds")
sub <- subset(TOP_30_GO_BP,  cluster_name == "MacroATP10A")
df <- subset(res@result, ID %in% sub$ID)

df <- as_tibble(df) %>% 
  arrange(.data$p.adjust) %>% 
  mutate(Description = stringr::str_to_title(.data$Description)) %>% 
  mutate(Description = fct_reorder(.data$Description, dplyr::desc(.data$p.adjust)))

ggplot(df, mapping = aes(x = .data$Description, y = -log10(.data$p.adjust))) + 
  geom_bar(aes(fill = .data$Count), stat = "identity") + 
  scale_fill_gradient("Gene Count", low = "grey", high = "brown") +
  coord_flip() + labs(title = "MacroATP10A", x ="") + theme_classic() +
  ylab(bquote("-log10(adj.P-val)")) +
  theme(axis.text = element_text(size = 14, color = "black"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y  = element_text(size = 12, color = "black", face = "bold"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8, color = "black"),
        legend.position = c(0.3, 0.09),
        legend.key.height = unit(0.3, 'cm'),
        legend.key.width = unit(0.3, 'cm')) 

ggsave(paste0(out.dir, "MacroATP10A_top_30_go_bp.pdf"), useDingbats = F, width = 8, height = 8,  dpi = 320)


```

